Dynamics of Laterally Propagating Flames in X-ray Bursts. III. Comparing Hotspot Evolution in Two and Three Dimensions

### 2022-02-05 xrb3/paper.tex
more work on the paper

\begin{document}
%======================================================================
% Title
%======================================================================
\title{Dynamics of Laterally Propagating Flames in X-ray Bursts. III. Three-Dimensional Flames}

@@

Our 3D simulation used a base grid of $1024^2 \times 128$ zones with 2
levels of refinement, the first jumping by a factor of 4 and the next
by a factor of 2.  The domain has a size $(2.048\times
10^5~\mathrm{cm})^2 \times 2.56\times 10^4~\mathrm{cm}$, giving a fine
grid resolution of $25~\mathrm{cm}$.  The initial model and other
simulation parameters are identical to those used in
\citet{harpole:2021}.  We used static mesh refinement, to achieve
better load balancing, fully refining the atmosphere below a height of
$3600~\mathrm{cm}$.  These simulations used the 7-isotope He-burning
network used in \citet{eiden:2020}.  The simulations were run on the
OLCF Summit supercomputer, on 342 to 1366 nodes, with 6 NVIDIA V100
GPUs per node.  The entire computation was offloaded to GPUs following
the strategy described in \citet{castro_gpu}.

@@

>>>>

\begin{document}
%======================================================================
% Title
%======================================================================
\title{Dynamics of Laterally Propagating Flames in X-ray Bursts. III. Comparing Hotspot Evolution in Two and Three Dimensions}

@@

Our 3D simulation used a base grid of $1024^2 \times 128$ zones with 2
levels of refinement, the first jumping by a factor of 4 and the next
by a factor of 2.  The domain has a size $(2.048\times
10^5~\mathrm{cm})^2 \times 2.56\times 10^4~\mathrm{cm}$, giving a fine
grid resolution of $25~\mathrm{cm}$.  The initial model and other
simulation parameters are identical to those used in
\citet{harpole:2021}.  We used static mesh refinement, to achieve
better load balancing, fully refining the atmosphere below a height of
$3600~\mathrm{cm}$.  These simulations used the 7-isotope He-burning
network used in \citet{eiden:2020}.  The simulations were run on the
OLCF Summit supercomputer, on 342 to 1366 nodes, with 6 NVIDIA V100
GPUs per node.  The entire computation was offloaded to GPUs following
the strategy described in \citet{castro_gpu}.  Overall, about 370,000
node-hours were used for the calculation.  The full state of the
calculation was output only every 0.005 s, as each output file is 2.7 TB
in size (for single precision data).  We output a few fields more
frequently for visualization.  We ran a 2D simulation with the same
resolution and refinement strategy to compare with here.

@@

Because the 3D flame does not stay perfectly circular, it is difficult
to define the flame speed using the same method employed in
\citet{eiden:2020}.  Instead, we will look at the mass of the ash
material to assess how quickly the burning is taking place.
Figure~\ref{fig:mass_plot} shows the mass of the different species as
a function of time, for the 2D and 3D calculation.  The axisymmetric
2D calculation has slightly less volume than the 3D domain, since
rotating about the symmetry axis produces a cylindrical domain
inscribed in the 3D domain.  To compensate for this, we scale the
masses by the initial \isot{He}{4} mass.  This plot shows that the
masses of the heavy species , especially \isot{C}{12} and
\isot{Si}{28} grow quickly with time, but the 2D simulation appears to
burn faster than the 3D simulation.  This is because the assumption of
axisymmetry there does not allow for the complex structure we see in
the evolution of composition in the 3D flame.  This suggests that the
3D flame speed is slower than the 2D flame speed.

\begin{figure}[t]
\centering
\plotone{mass_plot}
\caption{\label{fig:mass_plot} Mass of species scaled to initial He mass as a function of time.  The 3D simulation is shown as the solid lines and the 2D simulation is shown as the dashed lines.}
\end{figure}
